---
title: "PA1_template.Rmd"
author: "Jerry"
date: "November 19, 2016"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    toc: yes
---
# This is the R Markdown File for the Reproducible Research Course Project 1

The first Step is Reading and Processing the data

1. Code for reading the dataset and/or Processing the data
```{r}
raw_data <- read.csv("activity.csv")
```

2.Histogram of the total number of steps taken
```{r}
stepsperday <-aggregate(steps~date,data=raw_data,sum,na.rm=TRUE)
hist(stepsperday$steps,breaks=10,main ="Total steps per day",col="green",xlab="Steps")
```

3. Mean and Median Number of steps taken each day
```{r}
data_mean <-mean(raw_data$steps,na.rm=TRUE)
data_median <-median(raw_data$steps,na.rm=TRUE)
```
The Mean number of steps taken each day is `r data_mean`.
The Median number of steps taken each day is `r data_median`.

4.Time Series Plot of the average number of steps taken

```{r,warning=FALSE}
steps_interval <-aggregate(steps~interval,data=raw_data,mean,na.rm=TRUE)
plot(steps_interval$interval,steps_interval$steps, type="l",main="Average Steps per 5 minute interval",xlab="Interval", ylab="Number of Steps")
```

5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}
maxsteps <-max(steps_interval$steps)
```
The Maximum number of steps in the 5 minute interval was `r maxsteps`

6.Code to describe and show a strategy for imputing missing data
&
7.Histogram of the total number of steps taken each day after missing values are imputed


a) Calculating the number of missing values

```{r}
miss_data <- sum(is.na(raw_data$steps))
```
The Number of missing data points is `r miss_data`

b) Replace the missing values by mean and creating the histogram
```{r}
processed_data <-raw_data
processed_data$steps[is.na(processed_data$steps)]<-median(raw_data$steps,na.rm=TRUE)
processed_data_day <-aggregate(steps~date,data=processed_data,sum,na.rm=TRUE)
hist(processed_data_day$steps,col=terrain.colors(11),main="Total Steps Per Day",xlab="Steps",ylab="Frequency")
```

c) calculating the median and mean
```{r}
data_mean_new <-mean(processed_data$steps,na.rm=TRUE)
data_median_new <-median(processed_data$steps,na.rm=TRUE)
```
The New Mean number of steps taken each day is `r data_mean_new`.
The New Median number of steps taken each day is `r data_median_new`.


8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(lattice)
processed_data$date <-as.Date(processed_data$date)
processed_data$dayname <-weekdays(processed_data$date)
processed_data$weekend <- as.factor(ifelse(processed_data$dayname == "Saturday" |
                      processed_data$dayname == "Sunday", "weekend", "weekday"))
data_map <- aggregate(steps ~interval+weekend,processed_data,mean)
xyplot(steps ~interval|factor(weekend),data=data_map,layout=c(1,2),type="l",col="purple")
```

